gij bent 2

### 2025-09-30 paper/paper.qmd
Finish conclusion

## Discussion and conclusion

This article aimed to investigate whether and how speakers deviate from general expectations. We focussed on the affordances these deviations, defined as individual variation, can bring for alternation research, as well as what implicit network structures can teach us about the spread of linguistic innovations. We looked into the choice between the conservative *gij zijt* and the innovative *gij bent* in Colloquial Belgian Dutch with a dataset based on `r num_tweets_filtered %>% formatn` tweets by `r num_users_filtered %>% formatn` authors. By having several attestations from the same authors, we were able to devise a "linguistic profile" for each author, quantifying their behaviour with regards to expectations based on more general factors like gender and dialect region. Through elastic net regression, a modern regression technique, deviations were found for `r non_eliminated_count %>% formatn` users in the dataset (`r removed_distribution_prop[["FALSE"]] %>% formatp` of authors). `r innovative_count %>% formatn` (`r non_zero_distribution_prop[["innovative"]] %>% formatp`) were innovative, while `r conservative_count %>% formatn` (`r non_zero_distribution_prop[["conservative"]] %>% formatp`) were conservative. We also found a regional pattern in the deviations, with the central Brabant region differing from the other Flemish dialects. The Brabant region correct towards the innovative form, the other regions correct towards the conservative form. We also found a geospatial pattern, spanning horizontally, which shows a potential future path for the diffusion of the innovative form.

The consequences of our findings are twofold. In a broad sense, our results show that individual variation exists, and in fact is abundant in our case. This illustrates that an important dimension of variation is likely lost in traditional regression models. The regional and geospatial patterns in our data show that if people deviate from the general patterns we expect, this deviation is not random. In a sense, it means that there is "second order variation", variation *within* the variation, so to say. The geospatial distribution also provides an opportunity to anticipate upcoming diffusion patterns. Of course, we should not treat these patterns as oracles, but they can be useful nonetheless.

The predictive nature of the geospatial distributions bring us to the network analysis of our study. Our results show that the socially influential users in our dataset are generally more innovative. This might indicate that the diffusion of the innovative *bent* variant might be currently in the "early adopters" stage. The innovative variant can therefore be assumed to diffuse more broadly in the future. Of course, it remains to be seen just how accurate the network measure is for accurately predicting the current stage in the diffusion of an innovation. A diachronic network study of a fully completed diffusion pattern could shed more light in this regard.

Finally, our study shows that CBD is still very much "in flux". While one cannot speak of a single, monolithic CBD variety, the geospatial findings suggest that a horizontal expansion of at least this Antwerp-Brabantian feature could be possible. This would go against previous findings arguing against the "Brabant expansion", but it is not unthinkable when one takes the perspective of individual variation. As was mentioned in @sec-introduction, individual linguistic behaviour is not uniformly progressive or conservative. An uneven diffusion of CBD features might find its roots in this non-uniform individual behaviour.

- [algemeen idee van de individuele variatie]
- [conclusies voor tussentaal]
- [idee dat je op deze manier ook kunt vaststellen in welke fase van de verspreiding een fenomeen zit]

>>>>

## Discussion

This article aimed to investigate the role of individual variation and network effects in language change through a study of an ongoing innovation in Colloquial Belgian Dutch (CBD). In this section, we will discuss to what extent we were able to associate linguistic behaviour to specific individuals and their place within a larger community.

First, we wanted to investigate to what extent individuals differ in their linguistic behaviour from the "general" behaviour predicted from their dialect area and gender. We looked into the choice between the conservative *gij zijt* and the innovative *gij bent* in Colloquial Belgian Dutch (CBD) with a dataset based on `r num_tweets_filtered %>% formatn` tweets by `r num_users_filtered %>% formatn` authors. By having several attestations from the same authors, we were able to devise a "linguistic profile" for each author, quantifying their behaviour with regards to expectations based on the aforementioned general factors like gender and dialect region. Through elastic net regression, a modern regression technique, deviations were found for `r non_eliminated_count %>% formatn` users in the dataset (`r removed_distribution_prop[["FALSE"]] %>% formatp` of authors). `r innovative_count %>% formatn` (`r non_zero_distribution_prop[["innovative"]] %>% formatp`) were more innovative than expected from their profile, while `r conservative_count %>% formatn` (`r non_zero_distribution_prop[["conservative"]] %>% formatp`) were more conservative than expected. We also found a regional pattern in the deviations, with the central Brabant region differing from the other Flemish dialects. The Brabant region correct towards the innovative form, the other regions correct towards the conservative form. We also found a geospatial pattern, spanning horizontally, which shows a potential future path for the diffusion of the innovative form.

The consequences of our findings are twofold. In general, our results show that individual variation exists, and in fact is abundant in our case. This illustrates that an important dimension of variation is likely lost in traditional regression models. On a more specific level, the regional and geospatial patterns in our data show that if people deviate from the general patterns we expect, this deviation is not random. In a sense, it means that there is "second order variation", variation *within* the variation, so to say. The geospatial distribution also provides an opportunity to anticipate upcoming diffusion patterns. Of course, we should not treat these patterns as oracles, but they can be useful nonetheless.

The predictive nature of the geospatial distributions bring us to the network component of our study. Our results show that the socially influential users in our dataset are generally more innovative. This might indicate that the diffusion of the innovative *bent* variant might be currently in the "early adopters" stage. The innovative variant can therefore be assumed to diffuse more broadly in the future. Of course, it remains to be seen just how accurate the network measure is for accurately predicting the current stage in the diffusion of an innovation. A diachronic network study of a fully completed diffusion pattern could shed more light in this regard. Still, it seems that network or other influence information about speakers could function as a "derivative function" in the S-curve, indicating what direction the innovation is heading.

Finally, our study shows that CBD as a variety is still very much "in flux". While one cannot speak of a single, monolithic CBD variety, the geospatial findings suggest that a horizontal expansion of at least this Antwerp-Brabantian feature could be possible. This would go against previous findings arguing against the "Brabant expansion", but it is not unthinkable when one takes the perspective of individual variation. As was mentioned in @sec-introduction, individual linguistic behaviour is not uniformly progressive or conservative. An uneven diffusion of CBD features might find its roots in this non-uniform individual behaviour. This shows that a focus on the individual, as well as the place of that individual within the larger community, could offer new perspectives on existing questions in studies of language change.
